---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r load}
library(dplyr)
data <- read.csv("activity.csv")
data$time.interval <- sapply(data$interval, function(e) { m <- e; h <- 0; if(e >= 100) { m <- e %% 100; h <- e %/% 100; }; sprintf("%02d:%02d", h, m) })
data$interval <- sapply(data$interval, function(e) { m <- e; h <- 0; if(e >= 100) { m <- e %% 100; h <- e %/% 100; }; h * 60 + m })
stats <- data %>% group_by(date) %>% summarise(total.steps = sum(steps, na.rm = TRUE), mean.steps = mean(steps, na.rm = TRUE))
```


## What is mean total number of steps taken per day?
```{r steps-per-day}
hist(stats$total.steps, main = "Total Steps Per Day", breaks = 10, ylab = "Steps", xlab = "Date")
summary(stats$total.steps)
```

## What is the average daily activity pattern?
```{r average-daily-pattern}
plot(data %>% group_by(interval) %>% summarise(interval.mean = mean(steps, na.rm = TRUE)), type = "l", main = "Average Daily 24-hour Activity Pattern", ylab = "Interval Mean", xlab = "Time of Day")
```


## Imputing missing values
There are `r sum(as.integer(is.na(data$steps)))` NAs out of `r nrow(data)` observations in the activity dataset.

```{r}
interval.median <- data %>% group_by(interval) %>% summarise(interval.median = median(steps, na.rm = TRUE))

imputed.data <- merge(data, interval.median)

imputed.data$steps[is.na(imputed.data$steps)] <- imputed.data$interval.median[is.na(imputed.data$steps)]
imputed.data$interval.median <- NULL
```

There are `r sum(as.integer(is.na(imputed.data$steps)))` NAs out of `r nrow(imputed.data)` observations in the imputed activity dataset.




## Are there differences in activity patterns between weekdays and weekends?
