---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r load}
library(dplyr)
data <- read.csv("activity.csv")
data$time.interval <- sapply(data$interval, function(e) { m <- e; h <- 0; if(e >= 100) { m <- e %% 100; h <- e %/% 100; }; sprintf("%02d:%02d", h, m) })
data$interval <- sapply(data$interval, function(e) { m <- e; h <- 0; if(e >= 100) { m <- e %% 100; h <- e %/% 100; }; h * 60 + m })
stats <- data %>% group_by(date) %>% summarise(total.steps = sum(steps, na.rm = TRUE), mean.steps = mean(steps, na.rm = TRUE))
```

## What is mean total number of steps taken per day?
```{r steps-per-day}
hist(stats$total.steps, main = "Total Steps Per Day", breaks = 10, ylab = "Steps", xlab = "Interval")
summary(stats$total.steps)
```

## What is the average daily activity pattern?
```{r average-daily-pattern}
plot(data %>% group_by(interval) %>% summarise(interval.mean = mean(steps, na.rm = TRUE)), type = "l", main = "Average Daily 24-hour Activity Pattern", ylab = "Interval Mean", xlab = "Time of Day")
```


## Imputing missing values
There are `r sum(as.integer(is.na(data$steps)))` NAs out of `r nrow(data)` observations in the activity dataset.

```{r impute-values}
interval.mean <- data %>% group_by(interval) %>% summarise(interval.mean = mean(steps, na.rm = TRUE))

imputed.data <- merge(data, interval.mean)

NAs <- is.na(imputed.data$steps)
imputed.data$steps[NAs] <- imputed.data$interval.mean[NAs]
imputed.data$interval.mean <- NULL
```

There are `r sum(as.integer(is.na(imputed.data$steps)))` NAs out of `r nrow(imputed.data)` observations in the imputed activity dataset.

```{r imputed-steps-per-day}
imputed.stats <- imputed.data %>% group_by(date) %>% summarise(total.steps = sum(steps, na.rm = TRUE), mean.steps = mean(steps, na.rm = TRUE))
hist(imputed.stats$total.steps, main = "Total Steps Per Day (imputed)", breaks = 10, ylab = "Steps", xlab = "Date")
summary(imputed.stats$total.steps)

```

## Are there differences in activity patterns between weekdays and weekends?
```{r weekdays-weekend}
library(timeDate)
weekend <- isWeekend(imputed.data$date)
#
# Meet requirement to add weekend weekday two-value factor to the
# dataset. Useful if using lattice to plot.
#
imputed.data$weekday.weekend <- weekend
imputed.data$weekday.weekend <- sapply(imputed.data$weekday.weekend, function(e) {
  if(e == TRUE) {
    as.factor("weekend")
  } else {
    as.factor("weekday")
  }
})
#
# Generate table. Instructions allow for us to use any plotting system. So am
# using standard plot and don't really need the factor column just created above.
#
par(mfrow = c(2, 1))
plot(data[weekend,] %>% group_by(interval) %>% summarise(interval.mean = mean(steps, na.rm = TRUE)), type = "l", main = "Weekend", ylab = "Number of Steps", xlab = "Interval")
plot(data[!weekend,] %>% group_by(interval) %>% summarise(interval.mean = mean(steps, na.rm = TRUE)), type = "l", main = "Weekday", ylab = "Number of Steps", xlab = "Interval")
```

